---
title: Analysis of Subscription Rates
author: Julian Winternheimer
date: '2022-01-24'
slug: []
categories: []
tags:
  - new buffer
---

In this analysis we'll analyze the rate at which new signups subscribe to paying plans on Buffer. We will also examine how long it takes for new users to subscribe.

## 28-Day Subscription Rate
First we'll look at subscription rates. For each weekly cohort of new users, we'll calculate the proportion that paid for a subscription _within 28 days of signing up_.

```{r include = FALSE, message = FALSE, warning = FALSE}
# load libraries
library(buffer)
library(dplyr)
library(ggplot2)
library(scales)
library(zoo)
library(CausalImpact)
```

```{r eval = FALSE, include = FALSE}
# define sql query
sql <- "
  select
    date_trunc(u.signup_at_date, week(monday)) as signup_week
    , date_trunc(u.signup_at_date, month) as signup_month
    , u.user_id
    , count(distinct s.id) as subs
  from dbt_buffer.buffer_users u
  left join dbt_buffer.stripe_paid_subscriptions s
    on u.user_id = s.account_id
    and timestamp_diff(s.first_paid_invoice_created_at, u.signup_at, day) between 0 and 27
  where u.signup_at between '2020-01-01' and '2021-12-27'
  group by 1,2,3
"

# query bigquery
customers <- bq_query(sql = sql)

# save data
saveRDS(customers, "subscription_rates.rds")
```

```{r warning = FALSE, message = FALSE, include = FALSE}
# load data
customers <- readRDS("subscription_rates.rds") %>% 
  mutate(converted = subs >= 1)
```

The plot below shows the proportion of weekly new users that susbcribed to a paid plan within 28 days of signing up. We can clearly see that the subscription rate increased significantly when New Buffer was launched in August of 2021.

```{r echo = FALSE, warning = FALSE, message = FALSE}
# group customers by signup week
by_week <- customers %>%
  group_by(signup_week, converted) %>% 
  summarise(users = n_distinct(user_id)) %>% 
  mutate(percent = users / sum(users)) %>% 
  filter(converted) %>% 
  ungroup

# plot conversion rates
by_week %>% 
  filter(signup_week != max(signup_week)) %>% 
  buffplot(aes(x = signup_week, y = percent)) +
  geom_line(size = 0.8) +
  geom_vline(xintercept = as.Date("2021-08-02"), color = "darkred", linetype = 2) +
  scale_y_continuous(labels = percent) +
  labs(title = "28-Day Subscription Rate",
       subtitle = "New Buffer Launch on 2 August 2021",
       x = "Signup Week", y = NULL)
```


The average 28-day subscription rate of users that signed up in the weeks prior to August 2, 2021 was **2.01%**. 

For users that signed up on or after August 9, 2021, the 28-day subscription rate was **2.34%**.

This represents a relative increase of around 16%. 


```{r warning = FALSE, message = FALSE, echo = FALSE}
# calculate avg rates before and after new buffer launch
avg_before <- by_week %>% 
  filter(signup_week < "2021-08-02") %>% 
  summarise(avg_rate = mean(percent)) %>% 
  pull()

avg_after <- by_week %>% 
  filter(signup_week >= "2021-08-09") %>% 
  summarise(avg_rate = mean(percent)) %>% 
  pull()

# add lines to graph
by_week %>% 
  filter(converted & signup_week != max(signup_week)) %>% 
  buffplot(aes(x = signup_week, y = percent)) +
  geom_line(size = 0.8) +
  geom_vline(xintercept = as.Date("2021-08-02"), color = "darkred", linetype = 2) +
  geom_smooth(data = subset(by_week, signup_week < "2021-08-02"), 
              method = lm,
              se = FALSE) +
  geom_smooth(data = subset(by_week, signup_week >= "2021-08-09"), 
              method = lm,
              se = FALSE) +
  scale_y_continuous(labels = percent) +
  labs(title = "28-Day Subscription Rate",
       subtitle = "New Buffer Launch on 2 August 2021",
       x = "Signup Week", y = NULL)
```

## Causal Impact Analysis
We can use a technique called Causal Impact to estimate the effect that launching New Buffer had on subscription rates.

```{r}
# get min and max dates
min_date <- "2019-12-30"
max_date <- "2021-12-13"

# create time series object
rate_ts <- zoo(by_week$percent, by_week$signup_week)

# specify the pre and post periods
pre.period <- as.Date(c(min_date, "2021-08-01"))
post.period <- as.Date(c("2021-08-16", max_date))

# run analysis
impact <- CausalImpact(rate_ts, pre.period, post.period, 
                       model.args = list(niter = 5000))

# plot impact
plot(impact, c("original", "pointwise")) +
  labs(x = "Week",
       title = "Impact on 28-Day Conversion Rate", 
       subtitle = "New Buffer Launch on August 2")  +
  scale_y_continuous(labels = percent) +
  theme_light(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(size = rel(1.5), face = "bold"),
        plot.subtitle = element_text(size = rel(1.1)),
        plot.caption = element_text(color = "#777777", vjust = 0),
        axis.title = element_text(size = rel(.9), hjust = 0.95, face = "italic"),
        panel.grid.major = element_line(size = rel(.1), color = "#000000"),
        panel.grid.minor = element_line(size = rel(.05), color = "#000000"),
        plot.background = element_rect(fill = '#F5F5F5', color = '#F5F5F5'),
        panel.background = element_rect(fill = '#F5F5F5', color = '#F5F5F5'))
```


The plot above and summary below tell us that the subscription rate was approximately 2.4% after New Buffer launched. In the absence of that launch event, we would have expected the subscription rate to be around 2.0%. In relative terms, we're looking at an increase of around 16-18%. 

The positive effect we've observed so far is statistically significant and unlikely to be due to random fluctuation.

```{r}
# show summary of causal impact
summary(impact)
```

